### 2021-06-24 article/article.tex
Article and documentation updates

\section{Abstract}
	\emph{spinsim} is a \emph{python} package that simulates spin half and spin one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda} \cite{lam_numba_2015}, which is an \emph{LLVM} (Low Level Virtual Machine) \cite{lattner_llvm_2004} compiler, and other optimisations, to allow for fast and accurate evaluation on \emph{Nvidia Cuda} \cite{nickolls_scalable_2008} compatible systems using GPU parallelisation.
	\emph{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin half or spin one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin one systems of nitrogen vacancy centres (NVCs) and BECs.

\section{Keywords}
Time dependent Schroedinger equation; Spin one; Spin half; Integrator; GPU; Solver; python; numba;

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing \cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals.
	The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum.
	The simplest spin system, spin half (ie spin quantum number of \(\frac12\), also referred to as a qubit), is quantised into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons.
	However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin one quantum system (ie spin quantum number of \(1\), also referred to as a quitrit), which is quantised into three quantum spin levels.

@@

	% \texttt{AtomicPy} \cite{morris_qcmonkatomicpy_2018}
	In the past, the spinor Bose Einstein condensate (spinor BEC) lab at Monash University used an in-house, \emph{cython} based script, on which this package is based, and standard differential equation solvers (such as \emph{Mathematica}'s function \texttt{NDSolve}) to solve the Schroedinger equation for quantum sensing spin systems.
	Spin one systems are sometimes approximated to spin half for a faster execution time, at the cost of modelling all effects of the system.
	However, these methods are not completely optimised for our use case, and therefore come with some issues.

@@

@@

		In the spin one case there is no standard basis of operators \(A_j\).
		Choices include the Gell Mann matrices, and multiple dipole-quadrupole bases \cite{hamley_spin-nematic_2012}\cite{di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra \(\mathfrak{su}(3)\), which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group \(SU(3)\)) in the spin one system.
		With this in mind, we choose to represent the Hamiltonian a linear combination of matrices from a 4-dimensional subspace of \(\mathfrak{su}(3)\), consisting of the spin matrices \(J_x, J_y\) and \(J_z\), and a single quadrupole operator \(Q = \frac13\mathrm{diag}(1, -2, 1)\),
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q.
		\end{align}

		Note that \(Q\) is proportional to \(Q_{zz}\) \cite{hamley_spin-nematic_2012} and \(Q_0\) \cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		These are the only operators necessary to simulate many spin one quantum systems, including quadratic Zeeman splitting described by the Breit-Rabi formula.
		\emph{spinsim} can also be configured to solve a general spin one system by setting the Hamiltonian to an arbitrary point in \(\mathfrak{su}(3)\), using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{ux}(t) U_1 + \omega_{uy}(t) U_2 + \omega_{vx}(t) V_1 + \omega_{vy}(t) V_2.
		\end{align}
		Here the added operators are those defined in \cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the \(U_1, U_2, V_1\) and \(V_2\) operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

@@

		% The matrices \(J_x, J_y, J_z\) are regular spin operators, and \(Q\) is a quadrupole operator. Note that \(Q\) is proportional to \(Q_{zz}\) as defined by Hamley et al \cite{hamley_spin-nematic_2012}, and \(Q_0\) as defined by Di et al \cite{di_dipolequadrupole_2010}.
		% INLINE Q DIAG. WHY WE'VE PICKED THESE FOUR. For ground state spin one atoms in a magnetic field, ... Breit Rabi ... . COULD BE READILY EXTENDED TO INCLUDE FURTHER ELEMENTS OR TO HIGHER SPINS.

@@

		Thus, to solve Equation \eqref{eq:schroedinger} for \(\psi(t)\) given a particular \(\psi(t_0)\), one only needs to solve Equation \eqref{eq:unitary_schroedinger} for \(\mathcal{U}(t, t_0)\).
		% Note that because \(\mathcal{U}(t, t_0)\) is unitary, the calculated value of \(\psi(t)\) using this method will not change in magnitude, conserving probability.
		Also since \(\mathcal{U}(t, t_0)\) is unitary, it can be written as a matrix exponential of a skew-Hermitian (also termed anti-Hermitian) matrix.
		If the Hamiltonian is constant, then this exponential is % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i H\cdot(t - t_0)).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series \cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for \(\mathcal{U}(t, t_0)\) is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The Dyson series gives an explicit expression for \(\mathcal{U}(t, t_0)\) in terms of multiple time integrals over nested time commutators of \(H(t')\). % Cite 2? recent papers
		While the Dyson series has been used numerically, it has only been so recently, in particular because once truncated, it is in general no longer unitary.

@@

		A Magnus series over a time step \(T\) will be guaranteed to converge only if the spectral norm of \(H\) is sufficiently small, \(\int_{t_0}^{t_0 + T}\mathrm{d}t_1\|H(t_1)\| < \pi\)\cite{blanes_magnus_2009}. Furthermore, each subsequent term in the expansion increases in complexity rapidly. For these reasons, the Magnus expansion can be used as a time stepping method rather than a single step to solve the complete system. Time stepping can be made from the fact that time evolution operators can be split into a product via
		\begin{align}
			\mathcal{U}(t, t_0) &= \mathcal{U}(t, t_{n - 1})\mathcal{U}(t_{n - 1}, t_{n - 2})\cdots\mathcal{U}(t_2, t_1)\mathcal{U}(t_1, t_0).\label{eq:product_of_time_evolution}
		\end{align}
		Each of the time evolution operators can be approximated by a Magnus expansion.

@@

		Many unitary time stepping techniques have been developed based on the Magnus expansion \cite{auer_magnus_2018}.
		Some of these techniques use Gauss-Legendre quadrature sampling to avoid the integration of commutators of the Hamiltonian, producing simple expressions that can be used for unitary time stepping.
		In particular, we used the commutator free 4 (CF4) from \cite{auer_magnus_2018} as our integration method.
		Suppose we wish to evaluate a short time step of \(\mathcal{U}(t_f + \mathrm{d}t, t_f)\) by taking a short step \(\mathrm{d}t\) from a fine sample time \(t_f\).
		Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

		\begin{align}
			t_1 &= t_f + \frac12 \mathrm{d}t\left(1 - \frac{1}{\sqrt{3}}\right)\textrm{, and}\\
			t_2 &= t_f + \frac12 \mathrm{d}t\left(1 + \frac{1}{\sqrt{3}}\right).
		\end{align}

		Then given the matrices

		\begin{align}
			\overline{H}_1 &= \mathrm{d}t\left(\frac{3 + 2\sqrt{3}}{12}H(t_1) + \frac{3 - 2\sqrt{3}}{12}H(t_2)\right)\textrm{, and}\label{eq:cf4_sample_1}\\
			\overline{H}_2 &= \mathrm{d}t\left(\frac{3 - 2\sqrt{3}}{12}H(t_1) + \frac{3 + 2\sqrt{3}}{12}H(t_2)\right),\label{eq:cf4_sample_2}
		\end{align}

		the unitary time step can be calculated using

		\begin{align}
			\mathcal{U}(t_f + \mathrm{d}t, t_f) = \exp(-i\overline{H}_2)\exp(-i\overline{H}_1).\label{eq:cf4_implementation}
		\end{align}

	\subsection{Exponentiator}
		From Equation \eqref{eq:cf4_implementation}, evaluating matrix exponentials is a core part of the integration algorithm.
		However, the \emph{spinsim} package uses the user provided field functions \(\omega(t)\) as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array.
		This allows for implementation of matrix exponentiation that takes advantage of this representation.
		For all implementations of exponentiation, the \emph{spinsim} computes the matrix exponential

		\begin{align}
			E(\omega) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
			&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		\end{align}

		In practice, \(\omega\) is the Lie algebra (spin and quadrature) basis representation of \(\overline{H_1}\) or \(\overline{H_2}\) defined in Equations \eqref{eq:cf4_sample_1} and \eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation \eqref{eq:cf4_implementation}.
		For spin half, the default exponentiator is in an analytic form.
		For spin one, an exponentiator based on the Lie Trotter product formula \cite{moler_nineteen_2003} is used instead.
		Explicitly, this formula is

		\begin{align}
			\exp\left( X + Y\right) &= \lim_{n\to\infty} \left(\exp\left(\frac{X}{n}\right) \exp\left(\frac{Y}{n}\right)\right)^n
		\end{align}

		Where \(X,Y\in\mathbb{C}^{N\times N}\) are matrix operators, and \(n\in\mathbb{N}\).
		The equality holds for without the limit for all \(n\) if the operators commute, like it does for real and complex numbers.
		This tells us that we can approximate a matrix exponential of a linear combination of spin operators by reducing the size of the coefficients of these operators, taking the analytically simple exponentials of each of the operators \(-iJ_x, -iJ_y, -iJ_z\) and \(-iQ\) separately, multiplying each individual result together, and then raising the combination \(T\) to the power of the original reduction.

		For our spin one case, the exponential \(E(\omega)\) can be approximated as, for large \(2^\tau\),
		\begin{align}
			E(\omega) =& \exp\left(-i\omega_x J_x - i\omega_y J_y - i\omega_z J_z - i\omega_q Q\right)\\
			=& \exp\left(2^{-\tau}\left(-i\omega_x J_x - i\omega_y J_y - i\omega_z J_z - i\omega_q Q\right)\right)^{2^\tau}\\
			\approx& \biggl(\exp\left(-i\frac12\left(2^{-\tau} \omega_z J_z + 2^{-\tau}\omega_q Q\right)\right)\nonumber\\
			&\cdot\exp\left(-i\left(2^{-\tau} \omega_\phi J_\phi\right)\right)\nonumber\\
			&\cdot\exp\left(-i\frac12\left(2^{-\tau} \omega_z J_z + 2^{-\tau} \omega_q Q\right)\right)\biggr)^{2^\tau}\\
			=& \begin{pmatrix}
				\left(\cos\left(\frac{\Phi}{2}\right) e^{-iz}e^{-iq}\right)^2 & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{-iz}e^{-i\phi} & -\left(\sin\left(\frac{\Phi}{2}\right)e^{iq}e^{-i\phi}\right)^2\\
				\frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{-iz}e^{i\phi} & \cos(\Phi)e^{i4q} & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{iz}e^{-i\phi}\\
				-\left(\sin\left(\frac{\Phi}{2}\right)e^{-iq}e^{i\phi}\right)^2 & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{iz}e^{i\phi} & \left(\cos\left(\frac{\Phi}{2}\right) e^{iz}e^{-iq}\right)^2
		% 		\left(\frac{\Gamma}{PZ}\right)^2 & \frac{SP}{Z\Phi} & -\left(\frac{\Sigma P}{\Phi}\right)^2 \\
		% 		\frac{SP\Phi}{Z} & CP^4 & \frac{SPZ}{\Phi} \\
		% 		-\left(\frac{\Sigma \Phi}{P}\right)^2 & \frac{SPZ\Phi}{1} & \left(\frac{\Gamma Z}{P}\right)^2
			\end{pmatrix}^{2^\tau}\\
			=& T^{2^\tau}.
		\end{align}

		Here \(z = 2^{-\tau}\frac{\omega_z}{2}\), \(z = 2^{-\tau}\frac{\omega_q}{6}\), \(\phi = 2^{-\tau}\sqrt{\omega_x^2 + \omega_y^2}\), and \(\phi = \mathrm{atan}2(\omega_y, \omega_x)\).
		Once \(T\) is calculated, it is then recursively squared \(\tau\) times to obtain \(E(\omega)\).
		In practice the calculations here are done by finding the differences from the matrices from the identity to avoid floating point cancellation errors from subtracting a small number from 1.
		The approach used for spin one exponentiation means that this particular exponentiator cannot solve arbitrary spin one quantum systems, as that would require the ability to exponentiate a point in the full, 8-dimensional Lie algebra of \(\mathfrak{su}(3)\), rather than just the 4-dimensional subspace spanned by the subalgebra \(\mathfrak{su}(2)\) spanned by \(\{J_x, J_y, J_z\}\), and the single quadratic operator \(Q\).
		An exponentiator which can exponentiate an element of the full algebra is included in the package for completeness, though this was not rigorously tested as our lab does not deal with physical systems that would require this functionality to work.

@@

	\subsection{Discretisation and parallelisation}
		Consider quantising time with
		\begin{align}
			t_k &= t_0 + \mathrm{D}t\cdot k,
		\end{align}

		where \(\mathrm{D}t\) is the time step of the time series (in contrast to \(\mathrm{d}t\), a smaller time step for integration).
		If we define

		\begin{align}
			\psi_k = \psi(t_k)\textrm{ and}\\
			\mathcal{U}_k = \mathcal{U}(t_{k}, t_{k-1}),
		\end{align}

		then the time series of states \(\psi_k\) and time evolution operators \(\mathcal{U}_k\) satisfies

		\begin{align}
			\psi_k &= \mathcal{U}_k\psi_{k-1}\label{eq:integration_compilation}
		\end{align}

@@

		% The CF4 method is used to calculate each individual \(u^k_l\). Let the fine sample time be given by \(t_f = l\mathrm{d}t + t_k\). Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by

>>>>

\section{Abstract}
	The \emph{spinsim} \emph{python} package simulates spin half and spin one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda}\cite{lam_numba_2015}, which is an \emph{LLVM} (Low Level Virtual Machine)\cite{lattner_llvm_2004} compiler for \emph{Nvidia Cuda}\cite{nickolls_scalable_2008} compatible systems using GPU parallelisation. 
	Along with other optimisations, this allows for speed improvements of up to four orders of magnitude while keeping staying just as accurate.
	\emph{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin half or spin one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin one systems of nitrogen vacancy centres (NVCs) and BECs.

\section{Keywords}
Time dependent Schr\"odinger equation; Quantum; Physics; Spin one; Spin half; Integrator; Exponentiator; GPU; Parallel computing; Solver; python; numba;

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing\cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals.
	The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum.
	The simplest spin system, spin half (ie spin quantum number of \(\frac12\), also referred to as a qubit), is quantised into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons.
	However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin one quantum system (ie spin quantum number of \(1\), also referred to as a quitrit), which is quantised into three quantum spin levels.

@@

	% \texttt{AtomicPy}\cite{morris_qcmonkatomicpy_2018}
	In the past, the spinor Bose Einstein condensate (spinor BEC) lab at Monash University used an in-house, \emph{cython} based script, on which this package is based, and standard differential equation solvers (such as \emph{Mathematica}'s function \texttt{NDSolve}) to solve the Schroedinger equation for quantum sensing spin systems.
	Spin one systems are sometimes approximated to spin half for a faster execution time, at the cost of modelling all effects of the system.
	However, these methods are not completely optimised for our use case, and therefore come with some issues.

@@

	Given the recent boom in machine learning, many research computers are now equipped with advanced graphics processing units (GPUs).
	Their many cores, which can range from hundreds to tens of thousands in number, allows them to run some highly parallel algorithms much faster than a central processing unit (CPU).
	Examples include calculating colours for many pixels on a screen in the titular graphics processing, or the weights in a large neural network.
	By finding ways to parallelise the problem of solving quantum spin systems, we can use the many cores of a GPU to our advantage in this context as well.

@@

		In the spin one case there is no standard basis of operators \(A_j\).
		Choices include the Gell Mann matrices, and multiple dipole-quadrupole bases\cite{hamley_spin-nematic_2012}\cite{di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra \(\mathfrak{su}(3)\), which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group \(SU(3)\)) in the spin one system.
		With this in mind, we choose to represent the Hamiltonian a linear combination of matrices from a 4-dimensional subspace of \(\mathfrak{su}(3)\), consisting of the spin matrices \(J_x, J_y\) and \(J_z\), and a single quadrupole operator \(Q = \frac13\mathrm{diag}(1, -2, 1)\),
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q.
		\end{align}

		Note that \(Q\) is proportional to \(Q_{zz}\)\cite{hamley_spin-nematic_2012} and \(Q_0\)\cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		These are the only operators necessary to simulate many spin one quantum systems, including quadratic Zeeman splitting described by the Breit-Rabi formula\cite{mockler_atomic_1961} important to experiments in our lab.
		The \emph{spinsim} simulator can also be configured to solve a general spin one system by setting the Hamiltonian to an arbitrary point in \(\mathfrak{su}(3)\), using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{u1}(t) U_1 + \omega_{u2}(t) U_2 + \omega_{v1}(t) V_1 + \omega_{v2}(t) V_2.
		\end{align}
		Here the added operators are those defined in\cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the \(U_1, U_2, V_1\) and \(V_2\) operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

@@

		% The matrices \(J_x, J_y, J_z\) are regular spin operators, and \(Q\) is a quadrupole operator. Note that \(Q\) is proportional to \(Q_{zz}\) as defined by Hamley et al\cite{hamley_spin-nematic_2012}, and \(Q_0\) as defined by Di et al\cite{di_dipolequadrupole_2010}.
		% INLINE Q DIAG. WHY WE'VE PICKED THESE FOUR. For ground state spin one atoms in a magnetic field, ... Breit Rabi ... . COULD BE READILY EXTENDED TO INCLUDE FURTHER ELEMENTS OR TO HIGHER SPINS.

@@

		Thus, to solve Equation \eqref{eq:schroedinger} for \(\psi(t)\) given a particular \(\psi(t_0)\), one only needs to solve Equation \eqref{eq:unitary_schroedinger} for \(\mathcal{U}(t, t_0)\).
		% Note that because \(\mathcal{U}(t, t_0)\) is unitary, the calculated value of \(\psi(t)\) using this method will not change in magnitude, conserving probability.
		Since \(\mathcal{U}(t, t_0)\) is unitary, it can be written as a matrix exponential of a skew-Hermitian (also termed anti-Hermitian) matrix.
		If the Hamiltonian is constant, then this exponential is % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i H\cdot(t - t_0)).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series\cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for \(\mathcal{U}(t, t_0)\) is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The Dyson series\cite{kalev_integral-free_2020} gives an explicit expression for \(\mathcal{U}(t, t_0)\) in terms of multiple time integrals over nested time commutators of \(H(t')\).
		While the Dyson series has been used numerically\cite{kalev_integral-free_2020}, it has only been so recently, in particular because once truncated, it is in general no longer unitary.

@@

		A Magnus series over a time step \(T\) will be guaranteed to converge only if \(\int_{t_0}^{t_0 + T}\mathrm{d}t_1\|H(t_1)\| < \pi\)\cite{blanes_magnus_2009}. Furthermore, each subsequent term in the expansion increases in complexity rapidly. For these reasons, the Magnus expansion is used as a time stepping method rather than a single step to solve the complete system. Time stepping can be made from the fact that time evolution operators can be split into a product via
		\begin{align}
			\mathcal{U}(t, t_0) &= \mathcal{U}(t, t_{n - 1})\mathcal{U}(t_{n - 1}, t_{n - 2})\cdots\mathcal{U}(t_2, t_1)\mathcal{U}(t_1, t_0).\label{eq:product_of_time_evolution}
		\end{align}
		Each of the time evolution operators can be approximated by a Magnus expansion.

@@

		Many unitary time stepping techniques have been developed based on the Magnus expansion\cite{auer_magnus_2018}.
		Some of these techniques use Gauss-Legendre quadrature sampling to avoid the integration of commutators of the Hamiltonian, producing simple expressions that can be used for unitary time stepping\cite{auer_magnus_2018}. % Suggest not obvious
		In particular, we use the commutator free, fourth order method (CF4) from Reference~\cite{auer_magnus_2018}.
		Suppose we wish to evaluate a time step of \(\mathcal{U}(t + \mathrm{d}t, t)\) using the CF4 method.
		The Hamiltonians are sampled at times
		\begin{align}
			t_1 &= t + \frac12 \left(1 - \frac{1}{\sqrt{3}}\right)\mathrm{d}t\quad\text{and}\\
			t_2 &= t + \frac12 \left(1 + \frac{1}{\sqrt{3}}\right)\mathrm{d}t,
		\end{align} % maybe not rm d
		based on the second order Gauss-Legendre quadrature
		\begin{align}
			\overline{H}_1 &= \frac{3 + 2\sqrt{3}}{12}H(t_1) + \frac{3 - 2\sqrt{3}}{12}H(t_2)\quad\text{and}\label{eq:cf4_sample_1}\\
			\overline{H}_2 &= \frac{3 - 2\sqrt{3}}{12}H(t_1) + \frac{3 + 2\sqrt{3}}{12}H(t_2).\label{eq:cf4_sample_2}
		\end{align}
		% \, \quad \qquad
		The time evolution operator is then approximated by
		\begin{align}
			\mathcal{U}(t + \mathrm{d}t, t) = \exp(-i\overline{H}_2\mathrm{d}t)\exp(-i\overline{H}_1\mathrm{d}t).\label{eq:cf4_implementation}
		\end{align}% \approx? Unitary? End with sentence

	\subsection{Exponentiator}
		From Equation \eqref{eq:cf4_implementation}, evaluating matrix exponentials is a core part of the integration algorithm.
		However, the \emph{spinsim} package uses the user provided field functions \(\omega(t)\) as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array.
		This allows for implementation of matrix exponentiation that takes advantage of this representation.
		For all implementations of exponentiation, the \emph{spinsim} computes the matrix exponential
		\begin{align}
			E(\omega) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
			&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		\end{align}
		In practice, \(\omega\) is the Lie algebra (spin and quadrature) basis representation of \(\overline{H_1}\) or \(\overline{H_2}\) defined in Equations \eqref{eq:cf4_sample_1} and \eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation \eqref{eq:cf4_implementation}.

		For spin half, the default exponentiator is in an analytic form.
		For spin one, an exponentiator based on the Lie Trotter product formula\cite{moler_nineteen_2003} is used instead.
		Explicitly, this formula is
		\begin{align}
			\exp\left( X + Y\right) &= \lim_{n\to\infty} \left(\exp\left(\frac{X}{n}\right) \exp\left(\frac{Y}{n}\right)\right)^n
		\end{align}
		Where \(X,Y\in\mathbb{C}^{N\times N}\) are matrix operators, and \(n\in\mathbb{N}\).
		The equality holds for without the limit for all \(n\) if the operators commute, like it does for real and complex numbers.
		This tells us that we can approximate a matrix exponential of a linear combination of spin operators by reducing the size of the coefficients of these operators, taking the analytically simple exponentials of each of the operators \(-iJ_x, -iJ_y, -iJ_z\) and \(-iQ\) separately, multiplying each individual result together, and then raising the combination \(T\) to the power of the original reduction.

		For our spin one case, the exponential \(E(\omega)\) can be approximated as, for large \(2^\tau\),
		\begin{align}
			E(\omega) =& \exp\left(-i\omega_x J_x - i\omega_y J_y - i\omega_z J_z - i\omega_q Q\right)\\
			=& \exp\left(2^{-\tau}\left(-i\omega_x J_x - i\omega_y J_y - i\omega_z J_z - i\omega_q Q\right)\right)^{2^\tau}\\
			\approx& \biggl(\exp\left(-i\frac12\left(2^{-\tau} \omega_z J_z + 2^{-\tau}\omega_q Q\right)\right)\nonumber\\
			&\cdot\exp\left(-i\left(2^{-\tau} \omega_\phi J_\phi\right)\right)\nonumber\\
			&\cdot\exp\left(-i\frac12\left(2^{-\tau} \omega_z J_z + 2^{-\tau} \omega_q Q\right)\right)\biggr)^{2^\tau}\\
			=& \begin{pmatrix}
				\left(\cos\left(\frac{\Phi}{2}\right) e^{-iz}e^{-iq}\right)^2 & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{-iz}e^{-i\phi} & -\left(\sin\left(\frac{\Phi}{2}\right)e^{iq}e^{-i\phi}\right)^2\\
				\frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{-iz}e^{i\phi} & \cos(\Phi)e^{i4q} & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{iz}e^{-i\phi}\\
				-\left(\sin\left(\frac{\Phi}{2}\right)e^{-iq}e^{i\phi}\right)^2 & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{iq}e^{iz}e^{i\phi} & \left(\cos\left(\frac{\Phi}{2}\right) e^{iz}e^{-iq}\right)^2
			\end{pmatrix}^{2^\tau}\\
			=& T^{2^\tau}.
		\end{align}
		% \begin{align}
		% 	E(\omega) =& \exp\biggl(-i\omega_x J_x - i\omega_y J_y - i\omega_z J_z - i\omega_q Q\\
		% 	&- i\omega_{u1} U_1 - i\omega_{u2} U_2 - i\omega_{v1} V_1 - i\omega_{v2} V_2\biggr)\\
		% 	& \exp\biggl(2^{-\tau}\biggl(-i\omega_x J_x - i\omega_y J_y - i\omega_z J_z - i\omega_q Q\\
		% 	&- i\omega_{u1} U_1 - i\omega_{u2} U_2 - i\omega_{v1} V_1 - i\omega_{v2} V_2\biggr)\biggr)^{2^\tau}\\
		% 	\approx& \biggl(\exp\left(-i2^{-\tau} \omega_\phi J_\phi\right)\exp\left(-i2^{-\tau} \omega_{u\phi} U_{u\phi}\right)\\
		% 	&\cdot\exp\left(-i2^{-\tau} \omega_{v\phi} V_{v\phi}\right)\exp\left(-i2^{-\tau} \omega_z J_z -i2^{-\tau} \omega_q Q \right)\biggr)^{2^\tau}\\
		% 	=& \biggl(\begin{pmatrix}
		% 		\cos^2\left(\frac{\Phi}{2}\right) & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{-i\phi} & -\left(\sin\left(\frac{\Phi}{2}\right)e^{-i\phi}\right)^2\\
		% 		\frac{-i}{\sqrt{2}} \sin(\Phi)e^{i\phi} & \cos\left(\Phi\right) & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{-i\phi}\\
		% 		-\left(\sin\left(\frac{\Phi}{2}\right)e^{i\phi}\right)^2 & \frac{-i}{\sqrt{2}} \sin(\Phi)e^{i\phi} & \cos^2\left(\frac{\Phi}{2}\right)
		% 	\end{pmatrix}\\
		% 	&\cdot \begin{pmatrix}
		% 		\cos\left(\Phi_u\right) & 0 & -i \sin\left(\Phi_u\right)e^{-i\phi_u}\\
		% 		0 & 1 & 0\\
		% 		-i \sin\left(\Phi_u\right)e^{i\phi_u} & 0 & \cos\left(\Phi_u\right)
		% 	\end{pmatrix}\\
		% 	&\cdot \begin{pmatrix}
		% 		\cos^2\left(\frac{\Phi_v}{2}\right) & \frac{-i}{\sqrt{2}} \sin(\Phi_v)e^{-i\phi_v} & \left(\sin\left(\frac{\Phi_v}{2}\right)e^{-i\phi_v}\right)^2\\
		% 		\frac{-i}{\sqrt{2}} \sin(\Phi_v)e^{i\phi_v} & \cos\left(\Phi_v\right) & \frac{i}{\sqrt{2}} \sin(\Phi_v)e^{-i\phi_v}\\
		% 		\left(\sin\left(\frac{\Phi_v}{2}\right)e^{i\phi_v}\right)^2 & \frac{i}{\sqrt{2}} \sin(\Phi_v)e^{i\phi_v} & \cos^2\left(\frac{\Phi_v}{2}\right)
		% 	\end{pmatrix}\\
		% 	&\cdot \begin{pmatrix}
		% 		e^{-iz - iq} & 0 & 0\\
		% 		0 & e^{i2q} & 0\\
		% 		0 & 0 & e^{iz - iq}
		% 	\end{pmatrix}\biggr)^{2^\tau}\\
		% 	=& T^{2^\tau}
		% \end{align}
		Here \(z = 2^{-\tau}\frac{\omega_z}{2}\), \(z = 2^{-\tau}\frac{\omega_q}{6}\), \(\phi = 2^{-\tau}\sqrt{\omega_x^2 + \omega_y^2}\), and \(\phi = \mathrm{atan}2(\omega_y, \omega_x)\).
		Once \(T\) is calculated, it is then recursively squared \(\tau\) times to obtain \(E(\omega)\).
		In practice the calculations here are done by finding the differences from the matrices from the identity to avoid floating point cancellation errors from subtracting a small number from 1.

		The approach used for spin one exponentiation means that this particular exponentiator cannot solve arbitrary spin one quantum systems, as that would require the ability to exponentiate a point in the full, 8-dimensional Lie algebra of \(\mathfrak{su}(3)\), rather than just the 4-dimensional subspace spanned by the subalgebra \(\mathfrak{su}(2)\) spanned by \(\{J_x, J_y, J_z\}\), and the single quadratic operator \(Q\).
		An exponentiator which can exponentiate an element of the full algebra is included in the package for completeness, though this was not rigorously tested as our lab does not deal with physical systems that would require this functionality to work.

@@

	\subsection{Discretisation and parallelisation}
		Consider quantising time with
		\begin{align}
			t_k &= t_0 + \mathrm{D}t\cdot k,
		\end{align}
		where \(\mathrm{D}t\) is the time step of the time series (in contrast to \(\mathrm{d}t\), a smaller time step for integration).
		If we define
		\begin{align}
			\psi_k = \psi(t_k)\textrm{ and}\\
			\mathcal{U}_k = \mathcal{U}(t_{k}, t_{k-1}),
		\end{align}
		then the time series of states \(\psi_k\) and time evolution operators \(\mathcal{U}_k\) satisfies
		\begin{align}
			\psi_k &= \mathcal{U}_k\psi_{k-1}\label{eq:integration_compilation}
		\end{align}

@@

		% The CF4 method is used to calculate each individual \(u^k_l\). Let the fine sample time be given by \(t_I = l\mathrm{d}t + t_k\). Then as part of the CF4 method, the field functions are sampled at particular times based on the second order Gauss-Legendre quadrature, given by
